---
title: "Practical Day 4"
format: html
---

##Question 1

```{r}
library(tidyverse)
library(nycflights13)

View(flights)
glimpse(flights)

```


## Question 2
```{r}

Distances<- flights %>% filter(month==1)%>% 
            group_by(carrier) %>%                    # group_by also handles finding distinct values 
            summarise(mean_distance=mean(distance, na.rm = TRUE),
            sd_distance = sd(distance, na.rm = TRUE)) %>% 
             arrange(mean_distance)
Distances
```

## Question 3
We get sd = NA when there is only one observation and we get sd = 0 when all the observations are equal.
```{r}
# Showing the chosen carriers sd and relevant information
carrier_data <- flights %>%
  filter(month == 1) %>%
  group_by(carrier) %>%
  summarise(
    num_flights = n(),
    unique_distances = n_distinct(distance),
    sd_dist = sd(distance)
  ) %>%
  # Filter for the cases specified:
  filter(num_flights == 1 | sd_dist == 0)

print(carrier_data)
```
Carrier OO has just one flight so its sd = NA. This is because n-1 is going to become 0 making the sd undefined.
The rest of the carriers show that they have just one unique value showing that the distances are all the same value
resulting in sd of 0.

## Question4
```{r}
df_wide<- flights %>% 
  group_by(month,carrier) %>% 
  summarise(average_dep_delay= mean(dep_delay, na.rm =TRUE),.groups = 'drop') %>% 
  pivot_wider(
    names_from = carrier ,
    values_from = average_dep_delay
)
print(df_wide)
```
## Question 5
```{r}
flight_prop<-flights %>% 
  summarise(proportion = mean(dep_delay>0 & arr_delay<=0, na.rm=TRUE))
flight_prop

```
## Question 6 
```{r}
dt_flights<- flights
dt_flights_route<-dt_flights %>% 
  group_by(origin,dest) %>%       #groups by routes
  filter(n_distinct(carrier)>1) %>%  #groups with more than 1 airline flies
  ungroup()
dt_flights_route

```
```{r}
delay_flights <- dt_flights_route %>%group_by(origin,dest,carrier) %>% summarise(average_arrival_delay= mean(arr_delay,na.rm = TRUE))  %>% left_join(airlines, by='carrier')
delay_flights
```
```{r}
#worst avg. arrival delay
worst_delay <- delay_flights %>%
  group_by(origin,dest) %>% 
  slice_max(average_arrival_delay,n=1,with_ties = FALSE )
worst_delay

# Find the BEST airline for each route
best_delay <- delay_flights %>%
  group_by(origin, dest) %>%
  slice_min(average_arrival_delay, n = 1, with_ties = FALSE)
```


```{r}
# Combine them and calculate the gap
route_gaps <- bind_rows(worst_per_route, best_per_route) %>%
  group_by(origin, dest) %>%
  summarise(
    max_delay = max(average_arrival_delay),
    min_delay = min(average_arrival_delay),
    gap = max_delay - min_delay,
    .groups = "drop"
  ) %>%
  slice_max(gap, n = 1)

route_gaps


worst_delay %>% left_join(best_delay) %>% group_by(origin,dest) %>% summarise(mutate(gap= max_delay))

```

## Question 7
```{r}

```



