---
title: "Practical Day 4"
format: html
---

## Question 1

```{r}
library(tidyverse)
library(nycflights13)

View(flights)
glimpse(flights)

```

## Question 2

```{r}

Distances<- flights %>% filter(month==1)%>% 
            group_by(carrier) %>%                    # group_by also handles finding distinct values 
            summarise(mean_distance=mean(distance, na.rm = TRUE),
            sd_distance = sd(distance, na.rm = TRUE)) %>% 
             arrange(mean_distance)
Distances
```

## Question 3

We get sd = NA when there is only one observation and we get sd = 0 when all the observations are equal.

```{r}
# Showing the chosen carriers sd and relevant information
carrier_data <- flights %>%
  filter(month == 1) %>%
  group_by(carrier) %>%
  summarise(
    num_flights = n(),
    unique_distances = n_distinct(distance),
    sd_dist = sd(distance)
  ) %>%
  # Filter for the cases specified:
  filter(num_flights == 1 | sd_dist == 0)

print(carrier_data)
```

Carrier OO has just one flight so its sd = NA. This is because n-1 is going to become 0 making the sd undefined. The rest of the carriers show that they have just one unique value showing that the distances are all the same value resulting in sd of 0.

## Question4

```{r}
df_wide<- flights %>% 
  group_by(month,carrier) %>% 
  summarise(average_dep_delay= mean(dep_delay, na.rm =TRUE),.groups = 'drop') %>% 
  pivot_wider(
    names_from = carrier ,
    values_from = average_dep_delay
)
print(df_wide)
```

## Question 5

```{r}
flight_prop<-flights %>% 
  summarise(proportion = mean(dep_delay>0 & arr_delay<=0, na.rm=TRUE))
flight_prop

```

## Question 6

```{r}
dt_flights<- flights
dt_flights_route<-dt_flights %>% 
  group_by(origin,dest) %>%       #groups by routes
  filter(n_distinct(carrier)>1) %>%  #groups with more than 1 airline flies
  ungroup()
dt_flights_route

```

```{r}
delay_flights <- dt_flights_route %>%
  group_by(origin,dest,carrier) %>% 
  summarise(average_arrival_delay= mean(arr_delay,na.rm = TRUE))%>% 
  left_join(airlines, by='carrier')
delay_flights
```

```{r}
#maximum avg. arrival delay for each route
worst_delay <- delay_flights %>%
  group_by(origin,dest) %>% 
  slice_max(average_arrival_delay,n=1,with_ties = FALSE )
worst_delay
```

```{r}
# minimum avg. delay time for each route
best_delay <- delay_flights %>%
  group_by(origin, dest) %>%
  slice_min(average_arrival_delay, n = 1, with_ties = FALSE)
best_delay
```

```{r}
# 
route_gaps <- bind_rows(worst_delay, best_delay) %>%
  group_by(origin, dest) %>%
  summarise(
    max_delay = max(average_arrival_delay),
    min_delay = min(average_arrival_delay),
    gap = max_delay - min_delay,
    .groups = "drop"
  ) %>%
  slice_max(gap, n = 1)

route_gaps
```

```{r}
gap <- worst_delay %>% 
  bind_rows(best_delay) %>% 
  group_by(origin, dest) %>% 
  summarise(
    max_delay = max(average_arrival_delay),
    min_delay = min(average_arrival_delay),
    gap = max_delay - min_delay,
    worst_airline = carrier[which.max(average_arrival_delay)],
    best_airline  = carrier[which.min(average_arrival_delay)],
    .groups = "drop"
  )
greatest_gap <-gap %>% slice_max(gap)
greatest_gap
```

```{r}
flights %>% filter(origin=='JFK', dest=='ATL')
```

Looking at the route data the huge difference could be caused by airline-specific operational factors such as scheduling, hub congestion, and fleet utilisation rather than route characteristics alone especially between 9E and DL.

## Question 7

```{r}
#| echo: true
#| message: false
#| warning: false
#| results: hide
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}
remotes::install_github("SATVILab/UtilsDataRSV")

```

```{r}
library(UtilsDataRSV)
unclean_data <- structure(list(id = c("id_1", "id_2", "id_3", "id_4", "id_5", 
"id_6", "id_7", "id_8", "id_9", "id_10", "id_11", "id_12", "id_13", 
"id_14", "id_15", "id_16", "id_17", "id_18", "id_19", "id_20", 
"id_21", "id_22", "id_23", "id_24", "id_25", "id_26", "id_27", 
"id_28", "id_29", "id_30", "id_31", "id_32", "id_33", "id_34", 
"id_35", "id_36", "id_37", "id_38", "id_39", "id_40", "id_41", 
"id_42", "id_43", "id_44", "id_45", "id_46", "id_47", "id_48", 
"id_49", "id_50"), age = c(50L, 34L, 70L, 33L, 22L, 61L, 69L, 
73L, 62L, 56L, 71L, 33L, 73L, 44L, 45L, 46L, 24L, 70L, 46L, 76L, 
47L, 76L, 28L, 48L, 54L, 27L, 45L, 26L, 61L, 28L, 38L, 55L, 33L, 
36L, 62L, 58L, 72L, 31L, 34L, 51L, 61L, 64L, 26L, 28L, 60L, 29L, 
42L, 46L, 79L, 72L), gender = c("male", "male", "male", "female", 
"female", "male", "female", "male", "male", "female", "female", 
"male", "male", "female", "male", "male", "male", "male", "female", 
"male", "male", "male", "male", "female", "femal", "male", "female", 
"female", "female", "female", "male", "female", "female", "female", 
"male", "male", "female", "male", "female", "female", "male", 
"female", "female", "male", "male", "female", "male", "male", 
"male", "female"), height = c(174.4, 197.7, 174.1, 194.5, NA, 
180.4, 170.5, 157.4, 196.8, 165.1, 153, 197.4, 186, 157.1, 177.5, 
197.7, 179.3, 170.2, 182.4, NA, 165.4, 161, 168.5, 199.2, 157.7, 
154.6, 157.1, 184.5, 181, 194.6, 183.6, 186.9, 176.1, 183, 191.1, 
189.3, 199, 172, 165.6, 170.5, 150.5, 159.2, 192.1, 161.6, 162, 
153.8, 162.3, 186.6, 192.4, 174.9), weight = c(69.4, 62.3, 55.6, 
69.5, 78.6, 60.8, 72.2, 60.9, 75.1, 67.7, 82.5, 68.7, 67.8, 76.7, 
87, 61.1, 70.6, 63.3, 81.5, 59.2, 93.2, 87.3, 83.4, 80.9, 68.6, 
76.5, 93.7, 79.1, 92, 65.6, 85.4, 63.3, 79.7, 74.1, 63.3, 78.2, 
95.7, 95.1, 63.7, 66.1, 99.3, 81, 96.9, 73.3, 70.3, 83, 57.6, 
78.6, 61.9, 98.1), blood_type = c("O", "A", "O", "O", "B", "AB", 
"O", "O", "O", "AB", "A", "O", "O", "O", "B", "A", "B", "AB", 
"O", "AB", "A", "AB", "O", "B", "A", "A", "B", "AB", "A", "B", 
"B", "A", "O", "O", "O", "B", "O", "A", "A", "B", "A", "O", "AB", 
"A", "A", "O", "O", "B", "A", "O"), disease_status = c("diseased", 
"healthy", "healthy", "healthy", "healthy", "healthy", "diseased", 
"healthy", "diseased", "Healthy", "diseased", "healthy", "diseased", 
"healthy", "diseased", "healthy", "healthy", "healthy", "healthy", 
"healthy", "healthy", "diseased", "healthy", "diseased", "healthy", 
"healthy", "healthy", "healthy", "diseased", "diseased", "healthy", 
"healthy", "healthy", "diseased", "diseased", "diseased", "healthy", 
"diseased", "healthy", "healthy", "healthy", "healthy", "healthy", 
"diseased", "diseased", "diseased", "healthy", "healthy", "diseased", 
"diseased"), cholesterol = c(228, 223, 213, 198, 166, 151, 195, 
199, 189, 196, 221, 156, 185, 230, 234, 174, 185, 236, 235, 180, 
165, 220, 160, 153, 250, 153, 184, 242, 212, 179, 224, 233, 181, 
199, 220, 214, 214, 248, 191, 162, 203, 173, 199, 187, 248, 189, 
173, 212, 164, 247), glucose = c(96, 78, 101, 119, 103, 91, 86, 
NA, 77, 80, 115, 85, 88, 109, NA, 71, 90, 94, 91, 87, 113, 93, 
97, 118, 109, 80, 85, 119, 99, 108, 89, 108, 97, 116, 79, 84, 
75, 81, 119, NA, 106, 109, 75, 82, 84, 75, 76, 120, 119, 77), 
    smoker = c("yes", "yes", "yes", "yes", "no", "yes", "no", 
    "yes", "no", "no", "no", "no", "no", "yes", "no", "yes", 
    "yes", "yes", "yes", "yes", "yes", "yes", "yes", "yes", "no", 
    "no", "yes", "yes", "yes", "no", "no", "yes", "no", "yes", 
    "no", "yes", "no", "yes", "yes", "yes", "no", "no", "yes", 
    "no", "no", "no", "no", "no", "no", "yes"), exercise = c("occasional", 
    "regular", "occasional", "regular", "none", "occasional", 
    "regular", "none", "occasional", "none", "occasional", "none", 
    "none", "regular", "occasional", "none", "regular", "regular", 
    "none", "occasional", "none", "occasional", "occasional", 
    "occasional", "regular", "occasional", "regular", "regular", 
    "regular", "occasional", "occasional", "none", "none", "regular", 
    "occasional", "occasional", "none", "none", "none", "none", 
    "occasional", "regular", "regular", "none", "regular", "occasional", 
    "occasional", "none", "occasional", "regular"), income = c(84820L, 
    81547L, 22588L, 72490L, 74533L, 25338L, 41469L, 57315L, 63629L, 
    88662L, 62615L, 56261L, 58499L, 82232L, 77584L, 77275L, 38468L, 
    54510L, 91326L, 78611L, 31402L, 29586L, 21441L, 58269L, 84173L, 
    88295L, 37940L, 43750L, 69750L, 92356L, 82518L, 91455L, 68866L, 
    51178L, 68275L, 27689L, 35418L, 81318L, 62405L, 86851L, 25654L, 
    47553L, 74474L, 51409L, 22607L, 55360L, 96351L, 21516L, 41927L, 
    55810L), education = c("master", "bachelor", "PhD", "master", 
    "bachelor", "highschool", "PhD", "highschool", "PhD", "PhD", 
    "bachelor", "highschool", "master", "bachelor", "PhD", "PhD", 
    "PhD", "bachelor", "master", "highschool", "PhD", "highschool", 
    "bachelor", "master", "highschool", "highschool", "master", 
    "master", "bachelor", "PhD", "highschool", "PhD", "master", 
    "master", "master", "PhD", "highschool", "master", "master", 
    "highschool", "bachelor", "highschool", "bachelor", "PhD", 
    "bachelor", "highschool", "master", "highschool", "bachelor", 
    "bachelor"), region = c("North", "South", "North", "West", 
    "North", "West", "South", "South", "West", "South", "West", 
    "South", "West", "East", "North", "West", "North", "North", 
    "West", "North", "East", "West", "South", "North", "North", 
    "East", "East", "North", "North", "West", "South", "West", 
    "West", "East", "West", "North", "West", "North", "East", 
    "North", "West", "South", "South", "East", "North", "West", 
    "West", "East", "North", "East"), marital_status = c("divorced", 
    "single", "divorced", "divorced", "divorced", "divorced", 
    "divorced", "married", "divorced", "married", "divorced", 
    "widowed", "married", "single", "widowed", "widowed", "single", 
    "divorced", "widowed", "widowed", "single", "married", "single", 
    "married", "widowed", "married", "single", "single", "widowed", 
    "married", "widowed", "divorced", "single", "married", "single", 
    "widowed", "widowed", "married", "widowed", "divorced", "married", 
    "married", "divorced", "single", "married", "widowed", "divorced", 
    "divorced", "single", "divorced")), row.names = c(NA, -50L
), class = c("tbl_df", "tbl", "data.frame"))
view_cols(unclean_data)
```

From the output we can see what inconsistencies exist in data like the typo of femal instead of female. We also see the nulls.
