---
title: "Practical 1"
format: html
---

## Question 1: Missing Values

Identify all rows in the `airquality` dataset that contain missing values.

```{r}
rowSums(is.na(airquality))                 # counts element wise NA variables
airquality[rowSums(is.na(airquality))>0,]  # shows rows with atleast 1 NA
```

## Question 2: Summary Statistics

Calculate the **mean, standard deviation, minimum, and maximum** for the `Temp` (temperature) and `Ozone` variables in the `airquality` dataset.

```{r}
#dataset contains null values so we must handle them too
mean(airquality$Temp, na.rm = TRUE)
mean(airquality$Ozone, na.rm = TRUE)
sd(airquality$Ozone, na.rm = TRUE)
sd(airquality$Temp, na.rm = TRUE)
min(airquality$Ozone, na.rm = TRUE)
min(airquality$Temp, na.rm = TRUE)
max(airquality$Ozone, na.rm = TRUE)
max(airquality$Temp, na.rm = TRUE)
```

## Question 3: Linear Regression via Matrices

1.  Compute the regression coefficients using the matrix formula.

```{r}
Y <- as.matrix(cars$dist)   
X<- cbind(1, cars$speed)

# solve(t(X) %*% X) 
beta_hat <- solve(t(X)%*%X)%*%t(X)%*%Y  ##linear regression
beta_hat 
```

## Question 4: Verify Regression Coefficients with `lm()`

Verify that the matrix-based coefficients match those obtained using R’s built-in `lm()` function

```{r}
lrm = lm(dist~speed, data= cars)
lrm$coefficients
```
